A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-11-15 documents/plm.tex
typo
the-method-who-shall-not-be-named

As mentioned above, in the context of this project,
  a model is something that can predict data given settings of some parameters,
  and an objective function that can be used to find best values for those parameters.
We prefer to work in a probabilistic mode in which the objective function can be justified
  in terms of a likelihood (probability for the data given the model)
  or a posterior probability density function (likelihood times a prior pdf).
The PLM treats each \Kepler\ data point $I_{mn}$ as being
  predictable from a linear combination of data points $I_{m'n}$,
  where $m'$ is from the set of non-overlapping pixels $m'\in\set{M}_m$.
\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + e_{mn}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n} + \sum_{k=1}^{K} b_{mnk}\,d_{kn}
\quad,
\end{eqnarray}
where $I^{\ast}_{mn}$ is the prediction (by the model) for data point $I_{mn}$,
  $e_{mn}$ is a noise contribution or residual away from the prediction,
  and the $a_{mnm'}$ and $b_{mnk}$ are the parameters (linear coefficients of the prediction).
The parameters $a_{mnm'}$ and $b_{mnk}$ have indices $m$ because
  they are different for every pixel $m$,
  and they will even be different for every time step $n$;
  that is, there will be a separate best-fit value for the parameters for every $I_{mn}$.
If we presume that the residuals away from the prediction are normally distributed with zero mean
  and known variance,
  likelihood optimization reduces to chi-squared minimization.
We add to the standard chi-squared definition a regularization term
  (equivalent to multiplying the likelihood by a Gaussian prior pdf)
  that penalizes large absolute values for the the coefficients $a_{mnm'}$:
\begin{eqnarray}
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mnn'}]^2}{\sigma^2_{mn'}}
                 + \sum_{m'\in\set{M}_m} \frac{a_{mnm'}^2}{\Sigma^2}
\\
I^{\ast}_{mnn'} &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n'} + \sum_{k=1}^{K} b_{mnk}\,d_{kn'}
\quad,
\end{eqnarray}
where the $\sigma^2_{mn'}$ are the (presumed known) individual-pixel noise variances,
  and $\Sigma^2$ sets the strength of the regularization (or width of the prior pdf).
Although the variances $\sigma^2_{mn'}$ are in principle different for every time step $n'$,
  for nearly-constant stars in good data, there is not much variation;
  in what follows, we treat these uncertainties as fixed at
  \todo{DFM: some mean value (we can't ignore this,
   because it ``plays off against'' the regularization)}
The regularization variance $\Sigma^2$ we set to 0.3 (\todo{DFM TAKE NOTE; that's about 10 for the parameter you usually use.})
  because we expect almost all the coefficients $a_{mnm'}$ to be less than unity in magnitude.
We do not regularize the meta-data coefficients $b_{mnk}$,
  because we want the model to make use of the (easily interpretable) meta-data coefficients as much as possible.

@@

We need a metric for comparing outputs.
For now lets try plotting the power spectrum of the SAP, PDC, and PLM photometry for each star.

So DWH is imagining multi-panel plots: SAP, PDC, and PLM vs time,
  plus a power-spectrum plot vs frequency.

@@

\clearpage
\begin{figure}
~ % DFM: replace this ~ with a figure
\caption{One quarter of photometric data for star XXX.
  The top panel shows the \Kepler\ SAP photometry.
  The next panel shows the \Kepler\ PDC photometry.
  The next shows the PDC photometry.
  The bottom panel shows the power spectra of the three lightcurves.\label{fig:example1}}
\end{figure}

>>>>

As mentioned above, in the context of this project,
  a model is something that can predict data given settings of some parameters,
  and an objective function that can be used to find best values for those parameters.
We prefer to work in a probabilistic mode in which the objective function can be justified
  in terms of a likelihood (probability for the data given the model)
  or a posterior probability density function (likelihood times a prior pdf).
The \name\ treats each \Kepler\ data point $I_{mn}$ as being
  predictable from a linear combination of data points $I_{m'n}$,
  where $m'$ is from the set of non-overlapping pixels $m'\in\set{M}_m$.
\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + e_{mn}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n} + \sum_{k=1}^{K} b_{mnk}\,d_{kn}
\quad,
\end{eqnarray}
where $I^{\ast}_{mn}$ is the prediction (by the model) for data point $I_{mn}$,
  $e_{mn}$ is a noise contribution or residual away from the prediction,
  and the $a_{mnm'}$ and $b_{mnk}$ are the parameters (linear coefficients of the prediction).
The parameters $a_{mnm'}$ and $b_{mnk}$ have indices $m$ because
  they are different for every pixel $m$,
  and they will even be different for every time step $n$;
  that is, there will be a separate best-fit value for the parameters for every $I_{mn}$.
If we presume that the residuals away from the prediction are normally distributed with zero mean
  and known variance,
  likelihood optimization reduces to chi-squared minimization.
We add to the standard chi-squared definition a regularization term
  (equivalent to multiplying the likelihood by a Gaussian prior pdf)
  that penalizes large absolute values for the the coefficients $a_{mnm'}$:
\begin{eqnarray}
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mnn'}]^2}{\sigma^2_{mn'}}
                 + \sum_{m'\in\set{M}_m} \frac{a_{mnm'}^2}{\Sigma^2}
\\
I^{\ast}_{mnn'} &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n'} + \sum_{k=1}^{K} b_{mnk}\,d_{kn'}
\quad,
\end{eqnarray}
where the $\sigma^2_{mn'}$ are the (presumed known) individual-pixel noise variances,
  and $\Sigma^2$ sets the strength of the regularization (or width of the prior pdf).
Although the variances $\sigma^2_{mn'}$ are in principle different for every time step $n'$,
  for nearly-constant stars in good data, there is not much variation;
  in what follows, we treat these uncertainties as fixed at
  \todo{DFM: some mean value (we can't ignore this,
   because it ``plays off against'' the regularization)}
The regularization variance $\Sigma^2$ we set to 0.3 (\todo{DFM TAKE NOTE; that's about 10 for the parameter you usually use.})
  because we expect almost all the coefficients $a_{mnm'}$ to be less than unity in magnitude.
We do not regularize the meta-data coefficients $b_{mnk}$,
  because we want the model to make use of the (easily interpretable) meta-data coefficients as much as possible.

@@

We need a metric for comparing outputs.
For now lets try plotting the power spectrum of the SAP, PDC, and \name\ photometry for each star.

So DWH is imagining multi-panel plots: SAP, PDC, and \name\ vs time,
  plus a power-spectrum plot vs frequency.

@@

\clearpage
\begin{figure}
~ % DFM: replace this ~ with a figure
\caption{One quarter of photometric data for star \todo{KIC XXX}.
  The top panel shows the \Kepler\ SAP photometry;
    the vertical axis is \todo{something like median-divided brightness}
    and the horizontal axis is \todo{something like KBJD}.
  The next panel shows the \Kepler\ PDC photometry.
  The next shows the \name\ photometry.
  The bottom panel shows the power spectra of the three lightcurves.\label{fig:example1}}
\end{figure}
